Shapes analysis for time series.

### 2024-05-14 NEURIPS24/paper.tex
update
up

% citep
% \section{TODO}
% %De ce que je comprends , les gens font surtout des variational auto encoder, ou du masking
% %Mentionner que la méthode proposé peut être utilisé avec des réseaux
% %\sam{Approche NTK, fédéré personalisation, projet avec Axel, Lien entre continuous normalizing flow}
% Plan intro: 
% \begin{itemize}
%     \item Representation of patterns in time series has several applications in biology, but halso for classification, deep learning finds features automaticly.
%      Taking or not taking into account time can be done with DTW. However, either the representation is the result of a black box and need post analysis to be understand,
%       which is difficult to apply statiscal tools on it, or the representation is desgned by hand to select relevant features, but the capacity of analysis is limited by the knowledge of the designer.
%      \item In shape analysis, representation of shapes for statistical analysis in a long time issue, which has been solved mainly by the LDDMM framework. Representing shapes as diffeomorpshims of a common shape. 
%      \item However, this tool has not been yet applied on time series, even if there is already some tools developped for oriented curves. 
%      \item In this work, we propose to apply LDDMM to time series and to show its interest to derive an interpretable and unsupervised representation of patterns in time series.
%      We proposed some tools developped in the shape analysis community to the researcher interested in machine learning and time series ones.
%      \item First, we introduce our assumptions regarding the dataset and the main problem of diffeomorpshim learning. Then, we show how to solve it with LDDMM .
%       We expose how to apply it to time series by selecting wisely a RKHS kernel, motivating it with a representation theorem.
%       experiments on synthetic data are presented to show that the parameter of representation are identifiable when the time seriess are not too sharp and the kernels parameter well tuned. 
%       Then, we show how to apply the method on an unsupervised case of moouse respeitaory dataset and a supervised one on ... .

@@

 Our goal is to analyze the inter-individual variability of a time series dataset, which is of prime interest in medicine and biology \cite{wang2016research,bar2012studying,germain2023unsupervised}.
  More specifically, we aim to find an unsupervised features representation method which encode the specificity of an individual compared to another. 
  In physiology, studying the different \textit{shapes} in a time series related to biological phenomena and their variations according to individual or pathology is common.
  However, a \textit{shape} has no clear definition; it is more an intuitive way to speak about the silhouette of a pattern in a time series. In this paper, we refer to as the shape of a time series, the graph of this signal.

>>>>

% citep
% \section{TODO}
% %De ce que je comprends , les gens font surtout des variational auto encoder, ou du masking
% %Mentionner que la méthode proposé peut être utilisé avec des réseaux
% %\sam{Approche NTK, fédéré personalisation, projet avec Axel, Lien entre continuous normalizing flow}
% Plan intro: 
% \begin{itemize}
%     \item Representation of patterns in time series has several applications in biology, but halso for classification, deep learning finds features automaticly.
%      Taking or not taking into account time can be done with DTW. However, either the representation is the result of a black box and need post analysis to be understand,
%       which is difficult to apply statiscal tools on it, or the representation is desgned by hand to select relevant features, but the capacity of analysis is limited by the knowledge of the designer.
%      \item In shape analysis, representation of shapes for statistical analysis in a long time issue, which has been solved mainly by the LDDMM framework. Representing shapes as diffeomorpshims of a common shape. 
%      \item However, this tool has not been yet applied on time series, even if there is already some tools developped for oriented curves. 
%      \item In this work, we propose to apply LDDMM to time series and to show its interest to derive an interpretable and unsupervised representation of patterns in time series.
%      We proposed some tools developped in the shape analysis community to the researcher interested in machine learning and time series ones.
%      \item First, we introduce our assumptions regarding the dataset and the main problem of diffeomorpshim learning. Then, we show how to solve it with LDDMM .
%       We expose how to apply it to time series by selecting wisely a RKHS kernel, motivating it with a representation theorem.
%       experiments on synthetic data are presented to show that the parameter of representation are identifiable when the time seriess are not too sharp and the kernels parameter well tuned. 
%       Then, we sh-ow how to apply the method on an unsupervised case of moouse respeitaory dataset and a supervised one on ... .

@@

 Our goal is to analyze the inter-individual variability of a time series dataset, which is of prime interest in medicine and biology \cite{guscelli2019importance,wang2016research,bar2012studying,germain2023unsupervised}.
  More specifically, we aim to find an unsupervised features representation method which encode the specificity of an individual compared to another. 
  In physiology, studying the different \textit{shapes} in a time series related to biological phenomena and their variations according to individual or pathology is common.
  However, a \textit{shape} has no clear definition; it is more an intuitive way to speak about the silhouette of a pattern in a time series. In this paper, we refer to as the shape of a time series, the graph of this signal.
